Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-30 files/Chapter3/femconvergence.tex
Chapter 3 final

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{FEM computation of eigenvalues and eigenfunctions}
\textcolor{red}{Change H to W?}

Let $H$ be a Hilbert space with the inner product $c(\cdot,\cdot)$ and induced norm $||\cdot||_{W}$. Let $V$ be a linear subspace of $H$, with inner product defined by the bilinear form $b(\cdot,\cdot)$. It is assumed that the bilinear form $b$ is symmetric and that the assumptions in Section \ref{sec:existence:ModalAnalysis} holds.

@@

The pair $(\lambda^n, \bar{u}_n)$ correspond to the pair $(\lambda^h_k, u^h_k)$ which is the solution of Problem $E^h$. It is necessary to some of the theory to make the connection.

@@

In section \ref{sec:estimating_the_eigenvalues} it is proved that the error $| \lambda_k^h - \lambda_k|$ is large when $k$ is large. It is for example possible that $|\lambda^h_1 - \lambda_1|$ is sufficiently small while $\lambda^h_k$ cannot even be considered an approximation of $\lambda_k$ when $\displaystyle k > \frac{1}{2}N_e$.

@@

\textcolor{red}{********* Hou voorlopig hier *************}
\subsubsection*{Properties of eigenvalues}
It is well known that the eigenvalues of a eigenvalue problem such as \eqref{GEVP} form an infinite sequence $(\lambda_i)$. Assume that the eigenvalues $\lambda_i$ can be ordered $\lambda_1 \leq \lambda_2 \leq \lambda_3 \leq ...$ and denote the corresponding eigenfunctions $u_i$.

Since any multiple of a eigenfunction is still an eigenfunction, the eigenfunctions can be normalized so that $||u_i||_{W} = 1$ for all $i$.

Fix $m \in \mathbb{N}$. Let $\left\{ \phi_k \in V | k = 1,2,...,m \right\}$ be a set of linear independent admissible basis functions. Define $S^h := \text{span}\left\{\phi_k \in V | k = 1,2,...,m\right\}$ so that $S^h$ is a finite dimensional subspace of V. \label{sym:natural} \label{sym:Sh1}

The ordering of the eigenvalues remains valid in a finite dimensional space. The eigenvalues in the space $S^h$ is denoted as $\lambda_i^h$ with corresponding normalized eigenfunction $u^h_i$ for $i = 1,2,...,m$.

Also any subspace of $S^h$ will also be a subspace of $V$. So the minmax principle applies and a lower bound for the eigenvalues can be found:
\begin{equation}
	\lambda_i \leq \lambda_i^h.
\end{equation}
\textcolor{red}{********* Hou voorlopig hier *************}

@@

\newtheorem{Prop_5}[Prop_1]{Proposition} 
\begin{Prop_5}
	\label{Prop_5}
	For any $\epsilon > 0$, there exists a $\delta >0$ such that for $h<\delta$,
	\begin{eqnarray*}
	||u^{*} - Pu^{*}||_{W} < \epsilon \ \textrm{ for each } u \in B_{m},\\
	||u - Pu||_{W} < \epsilon \ \textrm{ for each } u \in B_{m}.
	\end{eqnarray*}
\end{Prop_5}
\begin{proof}
	Consider a set of basis functions $\Phi \subset S^h$ such that $\phi^(k) = 0$ for each $\phi \in \Phi$.

	$Pu$ is the closest element in $S^h$ to $u$. In particular, $||u-Pu||_{W}\leq ||u-\Pi u||_{W}$, where $\Pi u$ is the interpolation of $u$ into $S^h$.

	\textcolor{red}{Need to fix reference to approximation theorem. Need to wait until section 3.5 is done. The approximation theorem is moved to the end of section 3.5}

	From the Approximation Theorem given at the beginning of section \ref{sec:est}, is follows that
	\begin{eqnarray*}
		||u - \Pi u||_{W} \leq Ch^{k}||u^{(k)}||_{W}.
	\end{eqnarray*}

	So for any $\epsilon > 0$, a $\delta$ can be found such that if $h<\delta$ then
	\begin{eqnarray*}
		||u - Pu||_{W} \leq ||u - \Pi u||_{W} \leq \epsilon.
	\end{eqnarray*}
	A similar argument for $||u^{*} - Pu^{*}||_{W}$ proves the theorem.
\end{proof}

\newtheorem{Lem_3}[Lem_1]{Lemma} 
\begin{Lem_3}
	\label{Lem_3}
	For any $\epsilon >0$ there exists a $\delta > 0$ such that
	\begin{eqnarray}
	\sigma_{m}^{h} < \epsilon \ \textrm{ if } \ h < \delta. \label{eq:lem3}
	\end{eqnarray}
\end{Lem_3}
\begin{proof}
	For any $\epsilon > 0$ there exists a $\delta > 0$ such that if $h<\delta$, then
	\begin{eqnarray*}
	||u-Pu||_{W} < \epsilon \ \textrm{ for each } \ u \in B_{m}.
	\end{eqnarray*}
	This result, together with Lemma \ref{Lem_2} and Proposition \ref{Prop_5} gives the result.
\end{proof}

@@

The convergence of the eigenvalues follows from \eqref{eq:lem3} and \eqref{eq:lem4}.

@@

\subsubsection{The approximation theorem}

Consider a interpolation operator $\Pi$.\label{sym:interpolation} This projection is linear, i.e.
\begin{align*}
	\Pi(f + g) & = \Pi f + \Pi g,\\
	\Pi(\alpha f) & = \alpha \Pi f \ \text{ for a constant } \alpha. 
\end{align*}

@@

For the requirements of this dissertation, this can be simplified with an assumption. The assumption is that the basis of $S^h$ consists of polynomials. With this assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||_{W}$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}||_{W} \leq C h^{k-m} || u^{(k)}||_{W} \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}

This will be referred to as the approximation theorem in this section.
\end{document}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{FEM computation of eigenvalues and eigenfunctions}
Let $W$ be a Hilbert space with the inner product $c(\cdot,\cdot)$ and induced norm $||\cdot||_{W}$. In \cite{SF73}, the authors use a Hilbert space $H$, with inner product $(\cdot, \cdot)$ and induced norm $||\cdot||$. We remain with the notation that is consistent with this dissertation. Let $V$ be a linear subspace of $W$, with inner product defined by the bilinear form $b(\cdot,\cdot)$. It is assumed that the bilinear form $b$ is symmetric and that the assumptions in Section \ref{sec:existence:ModalAnalysis} holds.

@@

The pair $(\lambda^n, \bar{u}_n)$ correspond to the pair $(\lambda^h_k, u^h_k)$ which is the solution of Problem $E^h$. It is necessary to understand some of the theory to make the connection.

@@

In section \ref{sec:estimating_the_eigenvalues} it is proved that the error $| \lambda_k^h - \lambda_k|$ is large when $k$ is large. It is for example possible that $|\lambda^h_1 - \lambda_1|$ is sufficiently small while $\lambda^h_k$ cannot even be considered as an approximation for $\lambda_k$ when $\displaystyle k > \frac{1}{2}N_e$.

@@

@@

% \newtheorem{Prop_5}[Prop_1]{Proposition} 
% \begin{Prop_5}
% 	\label{Prop_5}
% 	For any $\epsilon > 0$, there exists a $\delta >0$ such that for $h<\delta$,
% 	\begin{eqnarray*}
% 	||u^{*} - Pu^{*}||_{W} < \epsilon \ \textrm{ for each } u \in B_{m},\\
% 	||u - Pu||_{W} < \epsilon \ \textrm{ for each } u \in B_{m}.
% 	\end{eqnarray*}
% \end{Prop_5}
% \begin{proof}
% 	Consider a set of basis functions $\Phi \subset S^h$ such that $\phi^(k) = 0$ for each $\phi \in \Phi$.
\subsubsection*{Assumption}
For any $\epsilon > 0$ there exists a $\delta > 0$ such that if $h<\delta$, then
\begin{eqnarray*}
||u-Pu||_{W} < \epsilon \ \textrm{ for each } \ u \in B_{m}.
\end{eqnarray*}

\textbf{Remark:} $Pu$ is the closest element in $S^h$ to $u$. In particular, $||u-Pu||_{W}\leq ||u-\Pi u||_{W}$, where $\Pi u$ is the interpolant of $u$ in $S^h$. The operator $\Pi$ is treated in Section \ref{sec:approximation_theorem}.

\newtheorem{Lem_3}[Lem_1]{Lemma} 
\begin{Lem_3}
	\label{Lem_3}
	For any $\epsilon >0$ there exists a $\delta > 0$ such that
	\begin{eqnarray}
	\sigma_{m}^{h} < \epsilon \ \textrm{ if } \ h < \delta. \label{eq:lem3}
	\end{eqnarray}
\end{Lem_3}

Substitute the assumption into the estimate for $\sigma^h_m$ in Lemma \ref{Lem_2}.

@@

The convergence of the eigenvalues follows from \eqref{eq:lem3} and \eqref{eq:lem4}. An estimate of the error depends on an estimate for $u - \Pi u$, see Section \ref{sec:approximation_theorem}

@@

\section{The approximation theorem} \label{sec:approximation_theorem}
Consider a interpolation operator $\Pi$.\label{sym:interpolation} This projection is linear, i.e.
\begin{align*}
	\Pi(f + g) & = \Pi f + \Pi g,\\
	\Pi(\alpha f) & = \alpha \Pi f \ \text{ for a constant } \alpha. 
\end{align*}

@@

For the requirements of this dissertation, this can be simplified with an assumption. The assumption is that the basis of $S^h$ consists of polynomials. With this assumptions, the semi-norm $|u|_H$ is equal to the norm $||u||_{W}$. This approximation theorem can be rewritten as
\newtheorem{Interpolation_2}{Theorem}
\begin{Interpolation_2}
	Suppose there exists an integer $k$ such that for each element
	\begin{equation*}
		s(\Pi_e) + 1 \leq k \leq r(\Pi_e) +1.
	\end{equation*}
	Then there exists a constant $C$ such that for any $u \in C^k_+(I)$,
	\begin{equation*}
		||(\Pi u)^{(m)} - u^{(m)}||_{W} \leq C h^{k-m} || u^{(k)}||_{W} \ \text{ for } m = 0,1,...,k.
	\end{equation*}
\end{Interpolation_2}
\end{document}
